---
title: "AIDS Drug"
author: "Anjolie"
date: "3/1/2021"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
```

```{r echo= FALSE}
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
triUmp$treatment <- as.factor(triUmp$treatment)
triUmp$infected <- as.factor(triUmp$infected)
triUmp$LGBTQ <- as.factor(triUmp$LGBTQ)
triUmp$DrugUser <- as.factor(triUmp$DrugUser)
triUmp$gender <- as.factor(triUmp$gender)
triUmp <- triUmp %>%
  filter(gender == "F") 

#for all cases
triUmp_count <- matrix(c(210,147,5484,5927),ncol=2,byrow=TRUE)
colnames(triUmp_count) <- c("Vaccine","Placebo")
rownames(triUmp_count) <- c("AIDS","None")
triUmp_count <- as.table(triUmp_count)
#for no drugs
TriUmp_noDrugs_count <- matrix(c(210,133,5386,5627),ncol=2,byrow=TRUE)
colnames(TriUmp_noDrugs_count) <- c("Vaccine","Placebo")
rownames(TriUmp_noDrugs_count) <- c("AIDS","None")
TriUmp_noDrugs_count <- as.table(TriUmp_noDrugs_count)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative AIDS drug (triUmpsal) is under test in clinical trials. Considering only the females, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.  
$H_0$: On average, the drug will NOT be more effective (on females) than the placebo.  
$H_A$: On average, the drug will be more or less effective (on females) than the placebo.  

# Methods
After filtering for only females, a contingency table will be made comparing the treatment ("drug" vs "placebo") with infection ("aids" vs "none"). We will also be comparing infection with whether the case was a drug user (for recreational/ unrelated medical use, because it's a different variable than "treatment") or not. A chi- squared test (and other inferential tests) will also be used.  
<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results
```{r}
#main
barchartGC(~infected + treatment, data = triUmp)
barchartGC(~infected + treatment, data = triUmp, type = "percent")
```
  
By looking at the percent graph, you tell that AIDS and the treatment drug are correlated; this means the drug is giving the patients AIDS (or at least making them more susceptible to it)!  

```{r}
#unrelated drugs vs infection
barchartGC(~DrugUser + infected, data = triUmp)
barchartGC(~DrugUser + infected, data = triUmp, type = "percent")
```
  
It may seem, at first glance, that taking unrelated drugs reduces your chance of avoiding AIDS, but the percent chart shows that this isn't true (because it takes into account that there are more people who don't take drugs/medicine than those who do). Whether unrelated drugs are taken doesn't affect the likelihood of getting AIDS at all.  
<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

### Numerical Descriptive Results
```{r}
rowPerc(triUmp_count)
colPerc(triUmp_count)
```
  
The row percents show that the vaccine is slightly correlated with AIDS and that not getting AIDS is slightly correlated with the placebo, suggesting the vaccine is doing the opposite of what's intended.  
<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

## Inferential Results
```{r}
chisq.test(triUmp_count)
chisqtestGC(triUmp_count)
fisher.test(triUmp_count)
oddsRatio = 1.543902
Prob = oddsRatio / (oddsRatio + 1)
print(Prob)
```
  
The odds ratio is at 1.5439 (with 95% confidence, between 1.2404 and 1.9257), which is greater than 1 and thus proves the alternate hypothesis true, especially with such small p- values, the biggest one being 0.0001 from Pearson's tests (which is greater than 0.05 and is probably rounded up from 0.00007686 anyway). The p- value means that, if the null hypothesis is true, there's a ~0.01% chance the data will be exactly as it is.  
The risk ratio (calculated with the below equation) is 1.5239. The risk ratio and odds ratio both being over 1 mean that someone vaccinated is more likely to get AIDS than someone who didn't - specifically 1.5239 times as likely. Thus the vaccine is not only ineffective but destructive.  
  
RiskRatio = vaccinated risk / unvaccinated risk;  
When a = AIDS, Vaccine,  
b = None, Vaccine,  
c = AIDS, Placebo, and  
d = None, Placebo, then:  
Vaccinated risk = a / (a+b) and  
Unvaccinated risk = c / (c+d).  

A vaccine's efficacy can be calculated with this equation:  
VE = (1-RiskRatio) * 100%  
Thus, the efficacy of this vaccine is -52.39%.  
<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion
The null hypothesis - that the vaccine is NOT more or less effective, on females, than the placebo - was proven wrong. The alternate hypothesis is correct because the vaccine is clearly destructive (at least on females), giving them AIDS (or at least making them more vulnerable to it). The use of unrelated drugs doesn't affect this. Thus, this vaccine should not be used on the general population.  
<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
